Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-01 paper/disc20clustering.tex
Intro to the lower bound

\maciek{Needed? "An~algorithm is \emph{component respecting}
if it always keeps nodes that  belong to the same component collocated.
That is,
if the algorithm needs to move a node,
it moves the whole component containing the node."}
%We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
%This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
%such that frequently communicating nodes (from different clusters)
%are relocated to the same cluster before they inflict too much communication cost.
%A \emph{reconfiguration} (i.e.,\emph{ re-partitioning})
%is performed by migrating nodes between clusters,
%at the total cost of individual node migrations.
%The objective is to jointly minimize communication and migration costs while serving the sequence.

We say that a component or a ground set is \emph{trivial}, if it contains a single node.

@@

The adversary constructs ground sets in reaction to choices of a deterministic online algorithm.
Once we construct a ground set, it lasts until the end of the input sequence.
If the algorithm splits any pair of nodes of a ground set, the adversary issue requests to any pair of split nodes.

We construct the ground sets in a way that they always form a balanced partition of nodes.
This way, the optimal algorithm may move to the ground-set-respecting configuration, and serve all requests for free.
(This implies that the algorithm must eventually move to a ground-set-respecting configuration if it is competitive.)

\maciek{Finished here.}

We utilize the fact that the online algorithm does not know the components of the adversary, and it may split some components while evicting nodes.
On the other hand, $\OPT$ initially collocates all nodes of each component and never splits them again.

\maciek{Here, we would like to introduce insights about the proof, emphasize the conditions for the bound to work (such as $k\geq 3$), and maybe compare to previously known lower bounds. Also, introduce the notation and definitions that are needed.}

\maciek{Note that the lower bound works for any value of $\alpha$, and we just use the number of migrations}

\begin{theorem}
  The competitive ratio of any deterministic online algorithm for \OBRP{} is in $\Omega(k\cdot \ell)$ for any $k\geq 3$.
  \maciek{TODO: the exact bound, not asymptotic}

  \maciek{The bound $(k \cdot \ell - k - \ell) / 2$ can be easily improved to $(k \cdot \ell - k - 1) / 2$ by playing a simple caching game in the first cluster, instead of just giving away the $k-1$ "blocker" (the caching game would build it gradually)}
\end{theorem}

>>>>

%We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
%This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
%such that frequently communicating nodes (from different clusters)
%are relocated to the same cluster before they inflict too much communication cost.
%A \emph{reconfiguration} (i.e.,\emph{ re-partitioning})
%is performed by migrating nodes between clusters,
%at the total cost of individual node migrations.
%The objective is to jointly minimize communication and migration costs while serving the sequence.

We say that a component or a ground set is \emph{trivial}, if it contains a single node.
We say that a~configuration is \emph{component respecting}
if the nodes belonging to the same component are collocated.
Similarly, we define a \emph{ground-set-respecting} configuration.

\maciek{Eviction - introduce.}

@@

The adversary constructs ground sets in reaction to choices of a deterministic online algorithm.
The algorithm does not know the ground set before they are revealed, and it may split some of them while evicting nodes.

Once we construct a ground set, it lasts until the end of the input sequence.
If the algorithm splits any pair of nodes of a ground set, the adversary issue requests to any pair of split nodes.

We construct the ground sets in a way that they always form a balanced partition of nodes.
This way, the optimal algorithm moves to a~ground-set-respecting configuration, and serves all requests for free.
(This implies that the algorithm must eventually move to any ground-set-respecting configuration if it is competitive.)

The lower bound requires $k\geq 3$.
It is oblivious to the choice of the reconfiguration cost $\alpha$, and we compare the number of performed node exchanges.

\begin{theorem}
	\label{th:lowerbound}
	The competitive ratio of any deterministic online algorithm for \OBRP{} is  $(k \cdot \ell - k - \ell) / 2 = \Omega(k\cdot \ell)$ for any $k\geq 3$.
\end{theorem}
